---
title: "CLSA PA and Sleep Analysis and Output"
author: "RSF"
date: "08/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1) Set-Up

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse,Hmisc, knitr, pander, broom, haven, rms, nlme,lspline, lavaan, semTools, tableone, ggpubr, ggplot2, openxlsx, expss)

setwd("~/Desktop/Manuscripts/Ongoing Projects/CLSA - Sleep and PA")
comp<-read.xlsx("Comprehensive_PA_Sleep_Processed.xlsx")#Processed Comprehensive Data (Complete Data) #26889 used same test language

#Select participants: exclude cognitively impaired, no PA data, and no sleep data

comp1<- subset(comp, Dementia.1==0 & Parkinsons.1==0 & SCI.1==0 & Stroke.1==0) #25830 without neurocognitive impairment

comp2<-subset(comp1, !is.na(PASE_TOTAL)) #20976 have PASE data
comp3<-subset(comp2, !is.na(SQ_Hours) & !is.na(SQ_Satisfaction) & !is.na(SQ_Latency_30min) & !is.na(SQ_Fragmentation) &
                !is.na(SQ_DaytimeSleepy) & !is.na(SQ_Apnea)) #20307 have sleep data


comp3$FAS_Total<-comp3$FAS_A_SCORE_COM + comp3$FAS_F_SCORE_COM + comp3$FAS_S_SCORE_COM


comp3<- comp3[c(1:3,170:213,247:259,4:169,260,214:260)]
```


2) Cognitive Function Score

```{r}
#Unitary Factor
Comp.Onefactor <- '

Global.Cog=~MAT_Score + RVLT_Immediate_Score + RVLT_Delayed_Score + Animal_Fluency_Strict + FAS_Total + PMT_TMT_Total + Stroop_Colour_Time + Stroop_Word_Time +Choice_RT_avgRT'

Onefit.comp <- cfa(Comp.Onefactor,data=comp3,missing="FIML")

pander(fitmeasures(Onefit.comp,c("chisq", "df", "cfi", 
                                 "tli","aic","bic","rmsea","rmsea.ci.lower",
                                 "rmsea.ci.upper")))


#Two Factor
Comp.Twofactor <- '
Memory=~RVLT_Immediate_Score + RVLT_Delayed_Score + PMT_TMT_Total

EF=~MAT_Score + Animal_Fluency_Strict + FAS_Total + Stroop_Colour_Time + Stroop_Word_Time + Choice_RT_avgRT'

Twofit.comp <- cfa(Comp.Twofactor,data=comp3,missing="FIML")

pander(fitmeasures(Twofit.comp,c("chisq", "df", "cfi", 
                                 "tli","aic","bic","rmsea","rmsea.ci.lower",
                                 "rmsea.ci.upper")))

#Three Factor
Comp.Threefactor <- '
Memory=~RVLT_Immediate_Score + RVLT_Delayed_Score + PMT_TMT_Total
Verbal=~Animal_Fluency_Strict+ FAS_Total
EF=~MAT_Score + Stroop_Colour_Time + Stroop_Word_Time +Choice_RT_avgRT'

Threefit.comp <- cfa(Comp.Threefactor,data=comp3,missing="FIML")

pander(fitmeasures(Threefit.comp,c("chisq", "df", "cfi", 
                                   "tli","aic","bic","rmsea","rmsea.ci.lower",
                                   "rmsea.ci.upper")))

#Compare Models
pander(anova(Twofit.comp,Onefit.comp))#Two Factor much better than 1
pander(anova(Threefit.comp,Onefit.comp))#Three Factor much better than 1
pander(anova(Threefit.comp,Twofit.comp))#Three Factor much better than 2


#standardized solution for three-factor measurement model
pander(standardizedsolution(Threefit.comp,ci=FALSE))

#Final Model
Comp.Final <- '
Memory=~RVLT_Immediate_Score + RVLT_Delayed_Score 
Verbal=~Animal_Fluency_Strict+ FAS_Total
EF=~MAT_Score + Stroop_Colour_Time + Stroop_Word_Time + Choice_RT_avgRT'

Finalfit.comp <- cfa(Comp.Final,data=comp3,missing="FIML")

pander(fitmeasures(Finalfit.comp,c("chisq", "df", "cfi", 
                                   "tli","aic","bic","rmsea","rmsea.ci.lower",
                                   "rmsea.ci.upper")))

pander(anova(Threefit.comp,Finalfit.comp))#Final better than full 3

#standardized solution for final measurement model
pander(standardizedsolution(Finalfit.comp,ci=FALSE))

#Final model latent scores for each participant
Latent_vars <- lavPredict(Finalfit.comp,append.data = FALSE)
comp4<- comp3
comp4$Memory<-Latent_vars[,1]
comp4$Verbal<-Latent_vars[,2]
comp4$EF<-Latent_vars[,3]
```

3) Descriptive Statistics

```{r}
dem_vars<-dput(names(comp4[c(3,2,4,11,20,31,6,50)]))
sleep_PA_vars<-dput(names(comp4[c(261,48,228,229,230,233,236,249)]))
Cog_vars<-dput(names(comp4[c(275:277)]))
Table1_dem_vars<-CreateTableOne(vars=dem_vars, data=comp4)
Table1_sleep_vars<-CreateTableOne(vars=sleep_PA_vars, data=comp4)
Table1_Cog_vars<-CreateTableOne(vars=Cog_vars, data=comp4)

#Demographic Variables#
print(Table1_dem_vars,contDigits=2,missing=TRUE,quote=TRUE)
#PA and Sleep Variables#
print(Table1_sleep_vars,contDigits=2,missing=TRUE,quote=TRUE)
#Cognitive Variable#
print(Table1_Cog_vars,contDigits=2,missing=TRUE,quote=TRUE)
```


4) Regression Models: Age associated changes in PA and Sleep Duration

```{r}
#data preparation
comp4$Age<-comp4$AGE_NMBR_COM
comp4$Sex<-NA
comp4$Sex[comp4$SEX_ASK_COM  == "M"]<-"Males"
comp4$Sex[comp4$SEX_ASK_COM  == "F"]<-"Females"

comp4$SQ_Satisfaction_1<-NA
comp4$SQ_Satisfaction_1[comp4$SQ_Satisfaction=="Very Dissatisfied"]<-1
comp4$SQ_Satisfaction_1[comp4$SQ_Satisfaction=="Dissatisfied"]<-2
comp4$SQ_Satisfaction_1[comp4$SQ_Satisfaction=="Neutral"]<-3
comp4$SQ_Satisfaction_1[comp4$SQ_Satisfaction=="Satisfied"]<-4
comp4$SQ_Satisfaction_1[comp4$SQ_Satisfaction=="Very Satisfied"]<-5

comp4$SQ_Fragmentation_1<-NA
comp4$SQ_Fragmentation_1[comp4$SQ_Fragmentation=="6-7x/wk"]<-1
comp4$SQ_Fragmentation_1[comp4$SQ_Fragmentation=="3-5x/wk"]<-2
comp4$SQ_Fragmentation_1[comp4$SQ_Fragmentation=="1-2x/wk"]<-3
comp4$SQ_Fragmentation_1[comp4$SQ_Fragmentation=="Less than 1x/wk"]<-4
comp4$SQ_Fragmentation_1[comp4$SQ_Fragmentation=="Never"]<-5

comp4$SQ_DaytimeSleepy_1<-NA
comp4$SQ_DaytimeSleepy_1[comp4$SQ_DaytimeSleepy=="6-7x/wk"]<-1
comp4$SQ_DaytimeSleepy_1[comp4$SQ_DaytimeSleepy=="3-5x/wk"]<-2
comp4$SQ_DaytimeSleepy_1[comp4$SQ_DaytimeSleepy=="1-2x/wk"]<-3
comp4$SQ_DaytimeSleepy_1[comp4$SQ_DaytimeSleepy=="Less than 1x/wk"]<-4
comp4$SQ_DaytimeSleepy_1[comp4$SQ_DaytimeSleepy=="Never"]<-5

comp4$SQ_Latency_30min_1<-NA
comp4$SQ_Latency_30min_1[comp4$SQ_Latency_30min=="6-7x/wk"]<-1
comp4$SQ_Latency_30min_1[comp4$SQ_Latency_30min=="3-5x/wk"]<-2
comp4$SQ_Latency_30min_1[comp4$SQ_Latency_30min=="1-2x/wk"]<-3
comp4$SQ_Latency_30min_1[comp4$SQ_Latency_30min=="Less than 1x/wk"]<-4
comp4$SQ_Latency_30min_1[comp4$SQ_Latency_30min=="Never"]<-5

comp4$SQ_OSA<-NA
comp4$SQ_OSA[comp4$SQ_Apnea=="Yes"]<-1
comp4$SQ_OSA[comp4$SQ_Apnea=="No"]<-0


compfinal<- comp4[-c(30,49,51,56:59,262,61:227,264)]
library(psych)

ddist<-datadist(compfinal)
options(datadist='ddist')
library(expss)
compfinal = apply_labels(compfinal,
                         Education.1 = "Education",
                         Avg_BMI = "BMI",
                         Smoking_Status.1 = "Smoking Status",
                         Lang_Used.1 = "Test Language",
                         Income_Level = "Income",
                         ADL.1 = "ADL",
                         PASE_TOTAL = "PASE Score",
                         SQ_Hours = "Sleep Duration Score"
)

#############PASE SCORE###################
PASE_Spline <- ols(PASE_TOTAL~rcs(Age, 3)*Sex + Avg_BMI +  Education4 + 
                     Income_Level + ADL + CESD_score,data=compfinal, x= TRUE, y = TRUE)

PASE_Spline
summary(PASE_Spline)#PASE score decreases with age (not significantly)
ggplot(Predict(PASE_Spline,Age,Sex), adj.subtitle = FALSE) + theme_classic2(base_size=14) + scale_colour_grey() + ylab("Total PASE Score")
pander(anova(PASE_Spline, tol= 1e-14))
plot(anova(PASE_Spline, tol= 1e-14, vnames="labels"),what="partial R2", xlim=c(0,.8), margin=c('partial R2', 'P'), cex = .9)



#############Sleep Duration###################
Duration_Spline <- ols(SQ_Hours~rcs(Age, 3)*Sex + Avg_BMI + Education4 + 
                         Income_Level + ADL + CESD_score,data=compfinal, x= TRUE, y = TRUE)

Duration_Spline
summary(Duration_Spline)#Sleep duration increases with age
ggplot(Predict(Duration_Spline,Age,Sex), adj.subtitle = FALSE) + theme_classic2(base_size=14) + scale_colour_hue() + geom_line(size=1) + ylab("Sleep Duration 
(hours/night)")
pander(anova(Duration_Spline))
plot(anova(Duration_Spline,vnames="labels"),what="proportion chisq", xlim=c(0,.8), margin=c('partial R2', 'P'), cex = .9)
```


5) Regression Models: Does PA moderate age-associated cognitive function?

```{r}
compfinal2<- compfinal%>%
  mutate(
    agebin=cut(Age, breaks=seq(from=45, to=86, by=5),right=FALSE, labels=FALSE),
  )%>%
  
  group_by(agebin,Sex)%>%
  mutate(
    PASE.norm = scale(PASE_TOTAL),
    Duration.norm = scale(SQ_Hours),
  )%>%
  ungroup()

ddist<-datadist(compfinal2)
options(datadist='ddist')
library(expss)

compfinal2 = apply_labels(compfinal2,
                          PASE.norm = "PASE Total Score",
                          Duration.norm = "Sleep Duration Score",
                          Quality1.norm = "Sleep Quality Score",
                          Quality2.norm = "Sleep Quality Score 2"
                          
)

#############PASE SCORE###################
#Memory
PASE_Memory <- ols(Memory~rcs(Age, 3)*Sex*rcs(PASE.norm, 3) + 
                     Avg_BMI + Education4 + 
                     Income_Level + ADL + Lang_Used + CESD_score, data=compfinal2)

PASE_Memory
ggplot(Predict(PASE_Memory,Age,PASE.norm=quantile(compfinal2$PASE.norm,na.rm=TRUE,c(.25,.75)), Sex), adj.subtitle = FALSE, linetype = PASE.norm) + theme_classic(base_size=14) + 
  scale_colour_manual(name="PASE Total Score",labels=c("25th %tile","75th %tile"),
                      values=c("#CCCCCC", "#000000"))
pander(anova(PASE_Memory, tol=1e-14))
plot(anova(PASE_Memory,vnames="labels", tol=1e-14),what="partial R2", xlim=c(0,.8), margin=c('partial R2', 'P'), cex = .9)



#Executive Function
PASE_EF <- ols(EF~rcs(Age, 3)*Sex*rcs(PASE.norm, 3) + 
                 Avg_BMI + Education4 + 
                 Income_Level + ADL + Lang_Used + CESD_score, data=compfinal2)



PASE_EF
ggplot(Predict(PASE_EF,Age,PASE.norm=quantile(compfinal2$PASE.norm,na.rm=TRUE,c(.25,.75)),Sex), adj.subtitle = FALSE, linetype = PASE.norm) + theme_classic(base_size=14) + 
  scale_colour_manual(name="PASE Total Score",labels=c("25th %tile","75th %tile"),
                      values=c("#CCCCCC", "#000000")) + ylab("Executive Function")
pander(anova(PASE_EF, tol=1e-14))
plot(anova(PASE_EF,vnames="labels", tol=1e-14),what="partial R2", xlim=c(0,.8), margin=c('partial R2', 'P'), cex = .9)



#Verbal Fluency
PASE_VF <- ols(Verbal~rcs(Age, 3)*Sex*rcs(PASE.norm, 3) + 
                 Avg_BMI + Education4 + 
                 Income_Level + ADL + Lang_Used + CESD_score, data=compfinal2)

PASE_VF
ggplot(Predict(PASE_VF,Age,PASE.norm=quantile(compfinal2$PASE.norm,na.rm=TRUE,c(.25,.75)),Sex), adj.subtitle = FALSE, linetype = PASE.norm) + theme_classic(base_size=14) + 
  scale_colour_manual(name="PASE Total Score",labels=c("25th %tile","75th %tile"),
                      values=c("#CCCCCC", "#666666", "#000000")) + ylab("Verbal Fluency")
pander(anova(PASE_VF, tol=1e-14))
plot(anova(PASE_VF,vnames="labels", tol=1e-14),what="partial R2", xlim=c(0,.8), margin=c('partial R2', 'P'), cex = .9)
```


6) Regression Models: Does sleep duration moderate age-associated cognitive function?

```{r}
#Memory
Duration_Memory <- ols(Memory~rcs(Age, 3)*Sex*rcs(Duration.norm, 3) + 
                         Avg_BMI + Education4 + 
                         Income_Level + ADL + Lang_Used + CESD_score, data=compfinal2)

Duration_Memory
ggplot(Predict(Duration_Memory,Age,Duration.norm=quantile(compfinal2$Duration.norm,na.rm=TRUE,c(.25,.75)),Sex), adj.subtitle = FALSE, linetype = Duration.norm) + theme_classic(base_size=14) + 
  scale_colour_manual(name="Sleep Duration",labels=c("25th %tile","75th %tile"),
                      values=c("#CCCCCC", "#000000"))
pander(anova(Duration_Memory, tol=1e-14))
plot(anova(Duration_Memory,vnames="labels", tol=1e-14),what="partial R2", xlim=c(0,.8), margin=c('partial R2', 'P'), cex = .9)



#Executive Function
Duration_EF <- ols(EF~rcs(Age, 3)*Sex*rcs(Duration.norm, 3) + 
                     Avg_BMI + Education4 + 
                     Income_Level + ADL + Lang_Used + CESD_score, data=compfinal2)

Duration_EF
ggplot(Predict(Duration_EF,Age,Duration.norm=quantile(compfinal2$Duration.norm,na.rm=TRUE,c(.25,.75)),Sex), adj.subtitle = FALSE, linetype = Duration.norm) + theme_classic(base_size=14) + 
  scale_colour_manual(name="Sleep Duration",labels=c("25th %tile","75th %tile"),
                      values=c("#CCCCCC", "#000000")) + ylab("Executive Function")
pander(anova(Duration_EF, tol=1e-14))
plot(anova(Duration_EF,vnames="labels", tol=1e-14),what="partial R2", xlim=c(0,.8), margin=c('partial R2', 'P'), cex = .9)



#Verbal Fluency
Duration_Verbal <- ols(Verbal~rcs(Age, 3)*Sex*rcs(Duration.norm,3) + 
                         Avg_BMI + Education4 + 
                         Income_Level + ADL + Lang_Used + CESD_score, data=compfinal2)


Duration_Verbal
ggplot(Predict(Duration_Verbal,Age,Duration.norm=quantile(compfinal2$Duration.norm,na.rm=TRUE,c(.25,.75)),Sex), adj.subtitle = FALSE, linetype = Duration.norm) + theme_classic(base_size=14) + 
  scale_colour_manual(name="Sleep Duration",labels=c("25th %tile","75th %tile"),
                      values=c("#CCCCCC", "#000000")) + ylab("Verbal Fluency")
pander(anova(Duration_Verbal, tol=1e-14))
plot(anova(Duration_Verbal,vnames="labels", tol=1e-14),what="partial R2", xlim=c(0,.8), margin=c('partial R2', 'P'), cex = .9)
```

7) Regression Models: Does PA predict age-associated variability in sleep duration?

```{r}
PASE_Duration <- ols(SQ_Hours~rcs(Age,3)*Sex*rcs(PASE.norm, 3) + 
                       Avg_BMI + Education4 + 
                       Income_Level + ADL + Lang_Used + CESD_score, data=compfinal2, x= TRUE, y = TRUE)


PASE_Duration
ggplot(Predict(PASE_Duration,Age,PASE.norm=quantile(compfinal2$PASE.norm,na.rm=TRUE,c(.25,.75)),Sex), adj.subtitle = FALSE, linetype = PASE.norm) + theme_classic(base_size=14) + 
  scale_colour_manual(name="PASE Score",labels=c("25th %tile","75th %tile"),
                      values=c("#CCCCCC", "#000000")) + ylab("Age Standardized Sleep Duration")
pander(anova(PASE_Duration, tol=1e-14))
plot(anova(PASE_Duration,vnames="labels", tol=1e-14),what="partial R2", xlim=c(0,.8), margin=c('partial R2', 'P'), cex = .9)
```

